---
title: "Red Wine Data Analysis"
author: "Madeline Li"
date: "6/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(leaps)
library(ggplot2)
library(reshape2)
library(MASS)
```

```{r}
dat = read.csv('winequality-red.csv', header = T)

facid <- dat$fixed.acidity
vacid <- dat$volatile.acidity
cacid <- dat$citric.acid
sugar <- dat$residual.sugar
chlor <- dat$chlorides
fsd <- dat$free.sulfur.dioxide
tsd <- dat$total.sulfur.dioxide
dens <- dat$density
ph <- dat$pH
sul <- dat$sulphates
qual <- as.factor(dat$quality)
alc <- dat$alcohol
```

```{r}
 get_lower_tri <- function(corr_white){
    corr_white[lower.tri(corr_white)]<- NA
    return(corr_white)
  }
```

```{r}
corr_white <- round(cor(dat),2)
melted_white_lower <- melt(get_lower_tri(corr_white))
```

```{r}
ggheatmap <- ggplot(melted_white_lower, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 3) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```

```{r}
pairs(dat)
cor(dat)
```

###Use Step-wise Regression with AIC
```{r}
mod0 = lm(alcohol ~ 1, data = dat)
mod.all = lm(alcohol ~., data = dat)
step(mod0, scope = list(lower = mod0, upper = mod.all))

fit <- lm(formula = alcohol ~ density + quality + fixed.acidity + pH + 
    residual.sugar + sulphates + free.sulfur.dioxide + citric.acid + 
    volatile.acidity + chlorides + total.sulfur.dioxide, data = dat)
```

### Check for Assumptions
```{r}
qqnorm(fit$residuals)
qqline(fit$residuals)
plot(fit$fitted.values,fit$residuals,
     xlab="Fitted Values",ylab="Residuals",main="Residuals vs Fitted")
abline(h = 0)
```

###Best Subsets Regression 
```{r}
library(leaps)
mod = regsubsets(cbind(facid, vacid, cacid, sugar, chlor, fsd, tsd, dens, ph, sul, qual), alc)
summary.mod = summary(mod)
summary.mod$which

summary.mod$rsq

summary.mod$adjr2

n = nrow(dat)
rss = summary.mod$rss
mse = c(rss[1]/(n-2), rss[2]/(n-3), rss[3]/(n-4), rss[4]/(n-5), rss[5]/(n-6), rss[6]/(n-7), rss[7]/(n-8), rss[8]/(n-9))
mse

summary.mod$cp
# facid, vacid, cacid, sugar, dens, ph, sul, qual

fit <- lm(alc ~facid+vacid+cacid+sugar+dens+ph+sul+qual)
```

### Check for Assumptions Again
```{r}
qqnorm(fit$residuals)
qqline(fit$residuals)
plot(fit$fitted.values,fit$residuals,
     xlab="Fitted Values",ylab="Residuals",main="Residuals vs Fitted")
abline(h = 0)
```


#Outliers
###Identify Outliers/influential points
```{r}
#Externally studentized finds influential points
#Internal is outlier
#Identify points that are outliers and identify points that are influential. Points that are both influential and outliers should be removed
outliers = abs(rstandard(fit))>3
influentials = abs(rstudent(fit))>3

for(x in 1:nrow(dat)){
  if(outliers[x] == TRUE) {
    if(influentials[x]==TRUE) {
      print(x)
    } 
  }
}

```

###Remove Points
```{r}
ndat = dat[-c(244, 245, 266, 354, 396, 468, 485, 494, 500, 557, 559, 560, 565, 634, 652, 654, 732, 1427, 1575)]
facid <- ndat$fixed.acidity
vacid <- ndat$volatile.acidity
cacid <- ndat$citric.acid
sugar <- ndat$residual.sugar
chlor <- ndat$chlorides
fsd <- ndat$free.sulfur.dioxide
tsd <- ndat$total.sulfur.dioxide
dens <- ndat$density
ph <- ndat$pH
sul <- ndat$sulphates
qual <- as.factor(ndat$quality)
alc <- ndat$alcohol
```

###Create new model without outliers/influential points
```{r}
#Best subsets
#R squared
library(leaps)
mod = regsubsets(cbind(facid, vacid, cacid, sugar, chlor, fsd, tsd, dens, ph, sul, qual), alc)
summary.mod = summary(mod)
summary.mod$which

summary.mod$rsq

summary.mod$adjr2

n = nrow(dat)
rss = summary.mod$rss
mse = c(rss[1]/(n-2), rss[2]/(n-3), rss[3]/(n-4), rss[4]/(n-5), rss[5]/(n-6), rss[6]/(n-7), rss[7]/(n-8), rss[8]/(n-9))
mse

summary.mod$cp
#Same model
fit <- lm(alc ~facid+vacid+cacid+sugar+dens+ph+sul+qual)
```

###Model Adequacy Checking Again
```{r}
plot(fit$fitted.values, fit$residuals,
     ylab="Residuals",
     xlab = "Fitted Values",
     main="Residuals vs Fitted Values")
abline(h=0,col="red")

qqnorm(fit$residuals)
qqline(fit$residuals)

```

###Research Questions

General linear F-test to test the two factors
```{r}
nullfit <- lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar +  pH + sulphates, data = dat) 
fullfit <- lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + density + pH + sulphates + quality, data = dat) 

anova(nullfit, fullfit)
```

Testing significance of all factors
```{r}
nullfit <- lm(alcohol ~ 1, data = dat)
fullfit <- lm(alcohol ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + density + pH + sulphates + quality, data = dat) 

anova(nullfit, fullfit)
```

Prediction Interval
```{r}
mfacid <- mean(ndat$fixed.acidity)
mvacid <- mean(ndat$volatile.acidity)
mcacid <- mean(ndat$citric.acid)
msugar <- mean(ndat$residual.sugar)
mchlor <- mean(ndat$chlorides)
mfsd <- mean(ndat$free.sulfur.dioxide)
mtsd <- mean(ndat$total.sulfur.dioxide)
mdens <- mean(ndat$density)
mph <- mean(ndat$pH)
msul <- mean(ndat$sulphates)
mqual <- round(mean(ndat$quality))

thedata = data.frame(facid = mfacid, vacid = mvacid, cacid = mcacid, sugar = msugar, chlor=mchlor, fsd = mfsd, tsd = mtsd, dens = mdens, ph = mph, sul = msul, qual = as.factor(mqual))

predict(fit, thedata, interval = "confidence", level = 0.95)
```










