---
title: "White Wine Analysis"
author: "Ishana Narayanan"
date: "June 5, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
white <- read.csv("winequality-white.csv")
```

```{r}
library(leaps)
library(ggplot2)
library(reshape2)
library(MASS)
library(car)
```

# Part 1: Data Basics

* fixed acidity = numeric, continuous
* volatile acidity = numeric, continuous
* citric acid = numeric, continuous
* residual sugar = numeric, continuous
* chlorides = numeric, continuous
* free sulfur dioxide = numeric, continuous
* total sulfur dioxide = numeric, continuous
* density = numeric, continuous
* pH = numeric, continuous
* sulphates = numeric, continuous
* alcohol = numeric, continuous
* quality = categorical, ordinal

# Part 2: Visualizing Data

```{r}
mod.full = lm(white$alcohol ~ white$fixed.acidity + white$volatile.acidity + white$citric.acid + white$residual.sugar + white$chlorides + white$free.sulfur.dioxide + white$total.sulfur.dioxide + white$density + white$pH + white$sulphates + white$quality)
```

(1) Scatterplot Matrix
```{r}
pairs(white, main="Scatterplot Matrix of White Wine Quality Data", pch=19)
```

(2) Correlation Matrix
```{r}
cor(white)
```

From the correlation matrix above, we see that there is a moderately positive linear relationship between alcohol and quality with a correlation value of 0.43557.

Significant negative linear relationships:

* alcohol and residual sugar (-0.45)

* alcohol and total sulfur dioxide (-0.448)

* alcohol and density (-0.78)

* pH and fixed acidity (-0.4258)

Significant positive linear relationships:

* residual sugar and density (0.83)

* density and total sulfur dioxide (0.529)

* residual sugar and total sulfur dioxide (0.401)

* total sulfur dioxide and free sulfur dioxide (0.6155)

(3) Correlation Heatmap
```{r}
corr_white <- round(cor(white),2)
melted_white <- melt(round(cor(white),2))
ggplot(melted_white, aes(x=Var1, y=Var2, fill=value)) + geom_tile()
```

```{r}
 get_lower_tri <- function(corr_white){
    corr_white[lower.tri(corr_white)]<- NA
    return(corr_white)
  }
```

```{r}
melted_white_lower <- melt(get_lower_tri(corr_white))
```

```{r}
ggheatmap <- ggplot(melted_white_lower, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 3) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```

# Part 3: Regression Analysis

## Best Subsets Regression

### Fitting Model

```{r}
mod = regsubsets(cbind(white$fixed.acidity, white$volatile.acidity, white$citric.acid, white$residual.sugar, white$chlorides, white$free.sulfur.dioxide, white$total.sulfur.dioxide, white$density, white$pH, white$sulphates, white$quality), white$alcohol)
summary.mod = summary(mod)
summary.mod$which
```

(1) $R^2$
```{r}
summary.mod$rsq
```


(2) Adjusted $R^2$
```{r}
summary.mod$adjr2
```

(3) Mallow's Cp Statistic
```{r}
summary.mod$cp
```

According to the adjusted $R^2$ and Mallow's Cp statistic, the best model is one with 8 predictors. Those predictors are fixed acidity, volatile acidity, residual sugar, free sulfur dioxide, density, pH, sulphates, and quality.

### Residual Analysis

```{r}
fit_subset = lm(white$alcohol ~ white$fixed.acidity + white$volatile.acidity + white$residual.sugar + white$free.sulfur.dioxide + white$density + white$pH + white$sulphates + white$quality)

plot(fit_subset)
```

According to the Residuals vs. Fit plot, we see that there are extreme outliers impacting these plots. We should remove the outliers so that we have a better understanding of what these plots are depicting.

### Removing Outliers

(1) Outliers

```{r}
out = rstudent(fit_subset)
which(abs(out) > 3)
```

(2) Influential Points

```{r}
hv <- hatvalues(fit_subset)
threshold <- 3*sum(hatvalues(fit_subset))/nrow(white)
which(abs(hv) > threshold)
```

(3) Cook's Distance

```{r}
cook <- cooks.distance(fit_subset)
plot(fit_subset, which=4)
```

```{r}
new_white <- white[-c(1527, 1654, 1664, 2782, 3902),]
```

### Returning to Residual Analysis

```{r}
fit_sub_no_outliers = lm(new_white$alcohol ~ new_white$fixed.acidity + new_white$volatile.acidity + new_white$residual.sugar + new_white$free.sulfur.dioxide + new_white$density + new_white$pH + new_white$sulphates + new_white$quality)
plot(fit_sub_no_outliers)
```

* In the Residuals vs. Fit plot, since the residuals are evenly spread about $x=0$ without any fanning, we can conclude that the error variances are equal. Moreover, this even spread of residuals, indicating that a linear model is a good fit.

* In the Normal Q-Q plot, the residuals are relatively following the curve. It is possible that this model is violating the non-normality LINE condition, so a transformation is needed to further investigate.

* There are still a few outliers, but not nearly as many in the previous Residuals vs. Fit plot.

* Thus, we can assume that from the above information that our model relatively meets the LINE conditions.

```{r}
summary(fit_sub_no_outliers)
```

## Stepwise Regression

### Build Regression Model

```{r}
mod0 = lm(white$alcohol ~ 1)
mod.full = lm(white$alcohol ~ white$fixed.acidity + white$volatile.acidity + white$citric.acid + white$residual.sugar + white$chlorides + white$free.sulfur.dioxide + white$total.sulfur.dioxide + white$density + white$pH + white$sulphates + white$quality)
step(mod0, scope=list(lower=mod0, upper=mod.full))
```

Model: alcohol ~ density + residual sugar + fixed acidity + pH + sulphates + volatile acidity + quality + free sulfur dioxide + citric acid + total sulfur dioxide

### Residual Analysis

```{r}
fit_step = lm(white$alcohol ~ white$density + white$residual.sugar + white$fixed.acidity + white$pH + white$sulphates + white$volatile.acidity + white$quality + white$free.sulfur.dioxide + white$citric.acid + white$total.sulfur.dioxide)
plot(fit_step)
```

According to the Residuals vs. Fit plot, we see that there are extreme outliers impacting these plots. We should remove the outliers so that we have a better understanding of what these plots are depicting.

### Removing Outliers

(1) Outliers

```{r}
out = rstudent(fit_step)
which(abs(out) > 3)
```

(2) Influential Points

```{r}
hv <- hatvalues(fit_step)
threshold <- 3*sum(hatvalues(fit_step))/nrow(white)
which(abs(hv) > threshold)
```

(3) Cook's Distance

```{r}
cook <- cooks.distance(fit_step)
plot(fit_step, which=4)
```

```{r}
new_white <- white[-c(1527, 1654, 1664, 2782, 3902),]
```

```{r}
fit_step_no_outliers = lm(new_white$alcohol ~ new_white$density + new_white$residual.sugar + new_white$fixed.acidity + new_white$pH + new_white$sulphates + new_white$volatile.acidity + new_white$quality + new_white$free.sulfur.dioxide + new_white$total.sulfur.dioxide)
plot(fit_step_no_outliers)
```

The residuals for the best subsets and stepwise regression models look relatively the same. Therefore, we should go with the model which has greater simplicity. Thus, the model we are going to further examine is the best subsets regression model.

Model: fit_subset = alcohol ~ fixed.acidity + volatile.acidity + residual.sugar + free.sulfur.dioxide + density + pH + sulphates + quality)

```{r}
summary(fit_step_no_outliers)
```

## Part 4: Transformations, Residual Analysis, and LINE Conditions

Since non-normality may potentially be an issue, we should take a log transformation on Y.

## 1. Log Transformation on Y

```{r}
fit_y = lm(log(new_white$alcohol) ~ new_white$fixed.acidity + new_white$volatile.acidity + new_white$residual.sugar + new_white$free.sulfur.dioxide + new_white$density + new_white$pH + new_white$sulphates + new_white$quality)
```

## Residual Analysis

```{r}
plot(fit_y)
```

* Residuals vs. Fit Plot

    + residuals are still evenly spread with no fanning, so the error terms have constant variance
  
    + linear fit is a good model
  
    + independent

* Normal Q-Q Plot

    + residuals follow the line better 
  
* LINE conditions are met

## 2. BoxCox Transformation

```{r}
trans=boxcox(fit_sub_no_outliers)

# (1) find the appropriate lambda based on max likelihood

# lambda based on max likelihood
lambda = trans$x[trans$y==max(trans$y)]
print(lambda)

# (2) apply transformation
new_alcohol = (new_white$alcohol)^lambda
```

lambda = 0.060606

```{r}
fit_boxcox = lm(new_alcohol ~ new_white$fixed.acidity + new_white$volatile.acidity + new_white$residual.sugar + new_white$free.sulfur.dioxide + new_white$density + new_white$pH + new_white$sulphates + new_white$quality)
plot(fit_boxcox)
```

* Residuals vs. Fit Plot

    + residuals are still evenly spread with no fanning, so the error terms have constant variance
  
    + linear fit is a good model
  
    + independent

* Normal Q-Q Plot

    + residuals follow the line better 
  
* LINE conditions are met


From the above transformations, we see that there are no significant changes with either a log transformation on Y or a BoxCox transformation. 

# Answering Research Questions

## Which factors affect alcohol content the most for white wine?

$H_0:$ $\beta_1$ = $\beta_2$ = $\beta_3$ = $\beta_4$ = $\beta_5$ = $\beta_6$ = $\beta_7$= $\beta_8$ = 0 

$H_a:$ at least one $\beta_k \neq 0$ for $k=1,2,3,4,5,6,7,8$

```{r}
mod.red = lm(new_white$alcohol ~ 1)
mod.full = fit_sub_no_outliers
anova(mod.red, mod.full)
```

Since the p-value of 2.2e-16 < $\alpha=0.05$, we must reject the null hypothesis. Therefore, not all $\beta_1 = \beta_2 = \beta_3 = \beta_4 = \beta_5 = \beta_6 = \beta_7= \beta_8 = 0$. There is sufficient evidence to conclude that some of the variation in alcohol content is explained by one or more of these predictors.

## Is the quality and density of wine significant in predicting the amount of alcohol in a particular wine?

$H_0:$ $\beta_5$ = $\beta_8$ = 0 

$H_a:$ at least one $\beta_k \neq 0$ for $k=5,8$

```{r}
mod.red = lm(new_white$alcohol ~ new_white$fixed.acidity + new_white$volatile.acidity + new_white$residual.sugar + new_white$free.sulfur.dioxide + new_white$pH + new_white$sulphates)
mod.full = fit_sub_no_outliers
anova(mod.red,mod.full)
```

Since the p-value of 2.2e-16 < $\alpha=0.05$, we must reject the null hypothesis. Therefore, both $\beta_5 = \beta_8 \neq 0$. There is sufficient evidence to conclude that quality and density of white wine is significant in predicting the amount of alcohol.

## What is the expected amount of alcohol in wine when all other factors are average?

```{r}
alcohol <- new_white$alcohol
fixed_acidity <- new_white$fixed.acidity
volatile_acidity <- new_white$volatile.acidity
residual_sugar <- new_white$residual.sugar
free_sulfur <- new_white$free.sulfur.dioxide
density <- new_white$density
pH <- new_white$pH
sulphates <- new_white$sulphates
quality <- new_white$quality
```

```{r}
fit = lm(alcohol ~ fixed_acidity + volatile_acidity + residual_sugar + free_sulfur + density + pH + sulphates + quality)

avg = data.frame(fixed_acidity = mean(fixed_acidity), volatile_acidity = mean(volatile_acidity), residual_sugar = mean(residual_sugar), free_sulfur = mean(free_sulfur), density = mean(density), pH = mean(pH), sulphates = mean(sulphates), quality = mean(quality))

predict(fit, avg, interval="confidence", level = .95)
```

We are 95% confident that the expected alcohol content is between 10.504 and 10.524 percent given that all other variables are averaged.